---
title: "Example"
author: "Laura Ospina-Rozo"
date: '2023-01-28'
output:
  html_document: 
    toc: yes
    toc_float: yes
    toc_depth: 5
    highlight: tango
    theme: paper
    self_contained: yes
    mode: selfcontained
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

This is an example to generate RMarkdown documents, and some types of figures and analysis that can be produced with R.

<br>
<br>

# Set up

## R and R studio 

These should be installed together. R console runs the code and R studio is the user interface.
<br>
In R studio, click "File/New project" and follow the prompts. It is good practice to have at least 4 folders: Data, Scripts, Output, Images. After that, click "File/New file/RMarkDown and then .
<br>


## Header

The first part of the RMarkdown code defines the properties of your web page. For example. `toc_depth: 5` means that there are 5 levels of subtitles displayed in the table of contents. `code_folding: hide` is to create buttons so that the reader can choose which code chunks they want to see. 

## Packages

Always include a list of the packages used in the code and even better to include the versions. Note I have already installed all these packages. And the code is used only to "activate them"

```{r message=FALSE, warning=FALSE}
# Import and process spectral data
library(pavo) 

# Plots
library(ggplot2)

# Tidy up
library(dplyr)
library(tidyr)

# To avoid running issues
library(here)

# To insert images
library(imager)
```

# Data wrangling

Use dplyr and tidyverse to arrange your data. 

mutate(): to create new columns, for example one column multiplied by another.

group_by() and summary(): When you have multiple measurements of the same sample and you want only the mean and sd. 

gather(): When you want to join two columns, one below the other. 

spread(): when you want to separate one column into two according to groups.

merge(): to join two data frames that have one column in common.

# Spectral Data

## Big files

```{r message = FALSE, warning=FALSE}

# Import data
BeetD1<-getspec(
  where = here::here("Data/Spectra"), # location
  ext = "txt", # extension of the files
  lim = c(300, 700), # limits
  decimal = ".")

BeetD2<-getspec(
  where = here::here("Data/Spectra"), # location
  ext = "txt", # extension of the files
  lim = c(450, 700), # limits
  decimal = ".")
```
*warning! models with this data are not reliable since we did not get values below 340nm


```{r}
# Change the names of the columns
# first column, wavelength should always be "wl"
names(BeetD2)<-c("wl","EpiLamp","CalFren","AnoAure","AnoLaeG")
```


```{r}
# Fixing data
BeetD1 <- BeetD1 %>%
  as.rspec() %>% # convert to an r spec file
  procspec(.,fixneg="zero") # Fix negatives

BeetD2 <- BeetD2 %>% 
  as.rspec() %>% # convert to an r spec file
  procspec(.,fixneg="zero") # Fix negatives
```

note: a smoothing function can also be applied if needed with "opt" and the numbers "span".

## Plots

```{r}
# Use this to explore the data
explorespec(BeetD2)

# Use this to produce a plot 
ggplot(data = BeetD2) +
  geom_line(aes(wl, AnoAure), col="blue")+
  geom_line(aes(wl, EpiLamp), col="green")+
  geom_line(aes(wl, CalFren), col="red")+
  xlab("Wavelength (nm)")+
  ylab("Reflectance (%)")+
  theme_bw()

# An extra rearrange of data is needed to automatically include a legend.

```
<br>
## Spectral parameters
<br>
Sometimes it is useful to identify the maximum wavelength (H1), or the area under the curve (B1). The function summary calculates these and some other parameters from the curves:

```{r}
summary(BeetD2)
```

Note: We advocate for the use of the spectral location (i.e. The wavelength at which reflectance is half of the maximum, nm ) instead of the maximum wavelength (sometimes called Hue).
<br>

For more details on these parameters and formulae to automate their extraction see:https://lospinarozo.github.io/OpticalPropertiesNaturalMaterials_RCode2021/#Introduction and http://bitly.ws/ARIM 

<br>
## Visual models
One last cool feature!

With pavo it is possible to plot spectra in a colour space
```{r}
vis.example <- 
  vismodel(BeetD1[,c(1,2)], # Spectra
           visual="cie10", # Organism /  Human
           illum = 'D65', # Illumination
           vonkries = TRUE, # corrections
           relative = FALSE, achromatic = 'none') # others

vis.cols <- colspace(vis.example, space='ciexyz')

plot(vis.cols, pch=21)

```

However, the parameters of the model may differ for each project, so this step requires more analysis.

For more details about the models see: https://mran.microsoft.com/snapshot/2017-02-04/web/packages/pavo/vignettes/pavo-1_0.html

# Microscopy Data 

This is the method I use to count layers.

## Tracing

This is the original image

```{r warning=FALSE, message=FALSE}
im1<-load.image(here::here("Images/Lats.png"))
plot(im1, axes = FALSE)
```

<br>

I use a vector graphics software to trace the layers and produce this: 

```{r}
im2<-load.image(here::here("Images/LatsLines.png"))
plot(im2, axes = FALSE)
```

And use plot profile in ImageJ to obtain the position of each line with reference to a starting point.

<br>

In this case, we will only consider one section of the png image. Import data.

```{r}
LatsP5L01<-read.csv(here::here("Data/Microscopy/A.laetus05LinesDataL1.csv")) 

```

This is what ImageJ produces:

```{r}
plot(LatsP5L01$Distance_.µm.,LatsP5L01$Gray_Value, type="l", xlab="Distance (um)", ylab="Gray value (0 to 255)")
```

## Extract

But from that data set, we can extract the distance between the points at which the reflectance is closer to 250 (max. gray value)

To do that, define a function to calculate the distance:

*Not all the vector images are the same, so we need to use different parameters for the grey value and minimal distance between layers. 

```{r}
#Define a function to extract the layers from the files
FindLayers2 <- function(Vector) {
  # filter values of distance with intensity > 245
  x<- filter(Vector, Gray_Value>245)
  # find difference between consecutive distance values
  xx <- diff(x$Distance_.µm.)[diff(x$Distance_.µm.)>0.0016] # adjust the  >xx distance for each spp:
  Lines <- xx
  Lines <- Lines*1000 # to nm since image measured in um
  return(Lines)
}
```

## Apply 

```{r}
LatsInn1 <- FindLayers2(LatsP5L01)[-1] # remove the first one because that is an artifact

plot(seq(1:length(LatsInn1)), # a sequence of numbers
     LatsInn1,# vector with the distances
     pch=23,col="black", bg="violet", # shapes and colour
     ylab="Number of the layer",  
     xlab= "Thickness (nm)" 
     )
```

<br>

# Stats

For example, I want to compare if the distance between the dots is the same in two beetles of the same species, different colour. In the following montage, the red beetle is on the left and the green is on the right. 

```{r warning=FALSE, message=FALSE}
im3<-load.image(here::here("Images/Montage.png"))
plot(im3, axes = FALSE)
```

I combine the plugin "Analyse particles" with "NND" in ImageJ to obtain the data on the distance between the dots. 

```{r}
#Import
AnSmaR <- read.csv(here::here("Data/Others/NNDAnoSmaR1.csv"))
AnSmaG <- read.csv(here::here("Data/Others/NNDAnoSmaG1.csv"))

NND<- c(AnSmaR$C1, AnSmaG$C1)
Morph<- c(rep("Red",length(AnSmaR$C1)),
          rep("Green",length(AnSmaG$C1)))

Smaragdinus <- data.frame(NND,Morph)

```

Plot

```{r}
ggplot(Smaragdinus, aes(Morph,NND))+
  geom_boxplot(bg="slateblue1", alpha=0.08, width=0.1)+
  geom_jitter(pch=16,col="slateblue3", alpha=0.4, cex=2, 
              width = 0.025)+
  theme_bw()+
  ylab("Nearest neighbour distance (um)")
```

t-test

```{r}
t.test(NND~Morph)
```

Interpretation: The centers of the hexagons are more separated in the green morph than in the red morph. 
